---
title: "cars"
author: "Yadu"
date: "July 23, 2015"
output: html_document
---
First we have to install the package ggplot2:
```{r}
#install.packages("ggplot2")
```

Next, we have to load the following packages (ggplot2 to generate scatterplots, plyr to change the column names, and datasets so that we can obtain the mtcars dataset):
```{r}
library(ggplot2)
library(plyr)
library(datasets)
```

Next, we change the names of the columns as follows:
```{r}
colnames(mtcars) <- c("Mileage (miles/gallon)", "Number of Cylinders", "Displacement (cubic inches)", "Gross Horsepower", "Rear Axle Ratio", "Weight (lb/1000)","Quarter Mile Time (Seconds)", "Engine Shape (V or Straight)", "Manual or Automatic Transmission", "Number of Forward Gears", "Number of Carburettors")
mtcars
```

Next, we have to substitute some of the quantitative data with qualitative data:
```{r}
mtcars$`Engine Shape (V or Straight)` <- gsub("0", "V", mtcars$`Engine Shape (V or Straight)`)
mtcars$`Engine Shape (V or Straight)` <- gsub("1", "Straight", mtcars$`Engine Shape (V or Straight)`)
mtcars$`Manual or Automatic Transmission` <- gsub("0", "Automatic", mtcars$`Manual or Automatic Transmission`)
mtcars$`Manual or Automatic Transmission` <- gsub("1", "Manual", mtcars$`Manual or Automatic Transmission`)
mtcars
```

Next, we obtain data regarding the internal structure of the data:
```{r}
str(mtcars)
summary(mtcars)
```

Next, we calculate the amount of time required to travel a full mile as follows:
```{r}
mtcars$`Full Mile Time (Seconds)` <- (mtcars$`Quarter Mile Time (Seconds)`)*4
```

The result is an entire new column consisting of quantitative data:
```{r}
mtcars$`Full Mile Time (Seconds)`
```

Next, we take the inverse of the quantity previously calculated to obtain the maximum speed of the vehicle:
```{r}
mtcars$`Speed (Miles Per Second)` <- 1/(mtcars$`Full Mile Time (Seconds)`)
```

Here also, we obtain an entire new column consisting of quantitative data:
```{r}
mtcars$`Speed (Miles Per Second)`
```

Next, we calculate the number of revolutions per minute as follows:
```{r}
mtcars$`Revolutions Per Minute` <- (mtcars$`Speed (Miles Per Second)`)*2112*(mtcars$`Rear Axle Ratio`)
```

Like the previous two data sets, this is also a column consisting of quantitative data:
```{r}
mtcars$`Revolutions Per Minute`
```

Next, we calculate torque as follows:
```{r}
mtcars$`Torque` <- ((mtcars$`Gross Horsepower`)*5252)/(mtcars$`Revolutions Per Minute`)
```

Again, we have another column consisting of quantitative data:
```{r}
mtcars$`Torque`
```

Next, we take two subsets of the newly updated dataset:
```{r}
cardata1 <- mtcars[c(15,4)]
cardata1
cardata2 <- mtcars[c(14,4)]
cardata2
```

We want to find out if horsepower have a linear correlation with any of the calculated quantities. We answer the question by generating scatterplots, lines of best fit, and loess curves of best fit as follows:
```{r}
require(ggplot2)

ggplot(data = cardata1, aes(x = cardata1$`Torque`, y = cardata1$`Gross Horsepower`))+
geom_point(shape = 1)+
geom_smooth(method = lm)+
xlab("Torque")+
ylab("Gross Horsepower")+
ggtitle("Correlation between Gross Horsepower and Torque")

ggplot(data = cardata1, aes(x = cardata1$`Torque`, y = 
cardata1$`Gross Horsepower`))+
geom_point(shape = 1)+
geom_smooth()+
xlab("Torque")+
ylab("Gross Horsepower")+
ggtitle("Correlation between Gross Horsepower and Torque")

ggplot(data = cardata2, aes(x = cardata2$`Revolutions Per Minute`, y = cardata2$`Gross Horsepower`))+
geom_point(shape = 1)+
geom_smooth(method = lm)+
xlab("Revolutions Per Minute")+
ylab("Gross Horsepower")+
ggtitle("Correlation between Gross Horsepower and Revolutions Per Minute")

ggplot(data = cardata2, aes(x = cardata2$`Revolutions Per Minute`, y = cardata2$`Gross Horsepower`))+
geom_point(shape = 1)+
geom_smooth()+
xlab("Revolutions Per Minute")+
ylab("Gross Horsepower")+
ggtitle("Correlation between Gross Horsepower and Revolutions Per Minute")
```

Then, we calculate the correlation coefficients for both sets of data as follows:

```{r}
cor(cardata1$`Gross Horsepower`, cardata1$`Torque`)
cor(cardata2$`Gross Horsepower`, cardata2$`Revolutions Per Minute`)

#Horsepower has an almost-linear correlation with torque but not with the number of revolutions per minute. A line of best fit does exist in the scatter plot. The correlation coefficient with torque is 0.91 which is close to 1 which indicates that the correlation is almost linear. The correlation coefficient with the number of revolutions per minute is 0.06 which is close to 0 which indicates that there is almost no linear correlation with the number of revolutions per minute. It is almost impossible for a line of best fit to exist. However, it is possible for a loess smothed fit curve to exist for both relations.     
```
